---
output:
  knitrBootstrap::bootstrap_document:
    theme.chooser: TRUE
    highlight.chooser: TRUE
---


Overlap with Neun+ peaks
================


Do differentially bound peaks found via `derfinder` overlap the published list of Neun+ peaks?

# Setup

```{r}
library('GenomicRanges')
library('devtools')
load('neunPeaks.Rdata')
runs <- c(paste0('run', 1:4, '-v1.0.10'), 'run5-v1.5.33')
ders <- lapply(runs, function(path) { 
    load(file.path('..', 'derAnalysis', path, 'fullRegions.Rdata'))
    return(fullRegions)
})
names(ders) <- runs
```


# Overlap: derfinder to peaks

Basic overlap info

```{r}
ov <- lapply(ders, function(run) {
    countOverlaps(run, neunPeaks)
})
ov.up <- lapply(ders, function(run) {
    countOverlaps(run, neunPeaks[neunPeaks$Direction == 'Up'])
})
ov.down <- lapply(ders, function(run) {
    countOverlaps(run, neunPeaks[neunPeaks$Direction == 'Down'])
})

## Does a DER overlap more than one peak?
lapply(ov, table)
#lapply(ov.up, table)
#lapply(ov.down, table)
```

The next subsections show the percent of candidate DERs overlapping known Neun+ peaks. First using all the peaks, next broken down by direction of the peak.

## All

```{r}
perc <- function(x) { sum(x > 0) / length(x) * 100 }


## Percent of candidate DERs overlapping known Neun+ peaks
sapply(ov, perc)

## Percent of significant DERs overlapping known Neun+ peaks
mapply(function(der, over) { perc(over[der$fwer < 0.05 ])}, ders, ov)
mapply(function(der, over) { perc(over[der$qvalues < 0.05 ])}, ders, ov)
```


## Up

```{r}
## Percent of candidate DERs overlapping known Neun+ peaks (Up only)
sapply(ov.up, perc)

## Percent of significant DERs overlapping known Neun+ peaks (Up only)
mapply(function(der, over) { perc(over[der$fwer < 0.05 ])}, ders, ov.up)
mapply(function(der, over) { perc(over[der$qvalues < 0.05 ])}, ders, ov.up)
```

## Down

```{r}
## Percent of candidate DERs overlapping known Neun+ peaks (Down only)
sapply(ov.down, perc)

## Percent of significant DERs overlapping known Neun+ peaks (Down only)
mapply(function(der, over) { perc(over[der$fwer < 0.05 ])}, ders, ov.down)
mapply(function(der, over) { perc(over[der$qvalues < 0.05 ])}, ders, ov.down)
```


# Overlap: Neun+ to derfinder

Do Neun+ peaks overlap regions found by the `derfinder` runs?


## All

```{r}
## Reverse direction: percent Neun+ peaks overlapping DERs, then Up only, then Down only
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks, der))
})
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks[neunPeaks$Direction == 'Up'], der))
})
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks[neunPeaks$Direction == 'Down'], der))
})
```


## FWER

```{r}
## Percent Neun+ peaks overlapping signicicant DERs (FWER), then Up only, then Down only
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks, der[der$fwer < 0.05 ]))
})
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks[neunPeaks$Direction == 'Up'], der[der$fwer < 0.05 ]))
})
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks[neunPeaks$Direction == 'Down'], der[der$fwer < 0.05 ]))
})
```


## FDR

```{r}
## Percent Neun+ peaks overlapping signicicant DERs (FDR), then Up only, then Down only
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks, der[der$fwer < 0.05 ]))
})
sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks[neunPeaks$Direction == 'Up'], der[der$fwer < 0.05 ]))
})

sapply(ders, function(der) {
    perc(countOverlaps(neunPeaks[neunPeaks$Direction == 'Down'], der[der$fwer < 0.05 ]))
})
```


# derfinder summary

Brief summary of number of regions found by `derfinder` and width distribution for the significant ones.

```{r}
## Length (in bp) summary of DERs
lapply(ders, function(der) { summary(width(der[der$fwer < 0.05 ]))})
lapply(ders, function(der) { summary(width(der[der$qvalues < 0.05 ]))})

## Number of significant DERs at 0.05
sapply(ders, function(der) { sum(der$fwer < 0.05) })
sapply(ders, function(der) { sum(der$qvalues < 0.05) })

## Total number of candidate DERs
sapply(ders, length)

## Percent of significant DERs at 0.05
sapply(ders, function(der) { sum(der$fwer < 0.05) }) / sapply(ders, length) * 100
sapply(ders, function(der) { sum(der$qvalues < 0.05) }) / sapply(ders, length) * 100
```



# Reproducibility

```{r}
## Reproducibility info
options(width = 120)
session_info()
Sys.time()
```


